Dominance Regions for Affine Cluster Algebras

### 2022-06-12 affine_dominance.tex
Modified our code

  The matrix $B$ is \emph{acyclic} if there is no sequence $i_1,\ldots,i_m,i_{m+1}=i_1$ so that $b_{i_\ell i_{\ell+1}}>0$ for $1\le\ell\le m$.
  This condition is naturally interpreted in terms of an associated quiver $Q_B$ with $n$ vertices and an arrow $i\to j$ whenever $b_{ji}>0$ (if we need to use representation theory at any point we'll have to change this to be $\gcd(b_{ij},b_{ji})$ arrows $i\to j$).
  Indeed, this is precisely the condition that $Q_B$ has no oriented cycles.
  In the case when $B$ is acyclic, there exists a labeling $\{k_1,\ldots,k_n\}=\{1,\ldots,n\}$ so that $\ell<\ell'$ implies $b_{k_{\ell'} k_\ell}\ge 0$.
  The sequence $(k_1,\ldots,k_n)$ (resp. $(k_n,\ldots,k_1)$) is then said to be \emph{source-adapted} (resp. \emph{sink-adapted}) since the condition implies $k_1$ is a source and $k_n$ is a sink in $Q_B$.
  %\sayDR{We need to pin down this convention.}

  For $b\in\ZZ$, write $[b]_+=max(b,0)$.
  Given a sign $\varepsilon\in\{\pm1\}$ and $k\in\{1,\ldots,n\}$, define a matrix $E_{k,\varepsilon}=(e_{ij})$ with
  \begin{equation}
    \label{eq:left mutation matrix}
    e_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ [\varepsilon b_{ik}]_+ & \text{if $i\ne j=k$;}\\ 0 & \text{otherwise;} \end{cases}
  \end{equation}
  and a matrix $F_{k,\varepsilon}=(f_{ij})$ with
  \begin{equation}
    \label{eq:right mutation matrix}
    f_{ij}=\begin{cases} 1 & \text{if $k\ne i=j$;}\\ -1 & \text{if $k=i=j$;}\\ [-\varepsilon b_{kj}]_+ & \text{if $k=i\ne j$;}\\ 0 & \text{otherwise.} \end{cases}
  \end{equation}
  Observe that $E^2_{k,\varepsilon}=\Id$ and $F^2_{k,\varepsilon}=\Id$ for any $k$ and any choice of $\varepsilon$.
  \begin{lemma}
    For $\varepsilon\in\{\pm1\}$ and $k\in\{1,\ldots,n\}$, we have $E_{k,-\varepsilon}E_{k,\varepsilon}=I_n+\varepsilon B^{\bullet k}$ and $E_{k,\varepsilon}-E_{k,-\varepsilon}=\varepsilon B^{\bullet k}$.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $\varepsilon b_{ij}=[\varepsilon b_{ij}]_+-[-\varepsilon b_{ij}]_+$.
  \end{proof}

  The index $k\in\{1,\ldots,n\}$ also determines a new skew-symmetrizable matrix $\mu_k B=(b'_{ij})$ given by
  \[ b'_{ij}=\begin{cases} -b_{ij} & \text{if $i=k$ or $j=k$;}\\ b_{ij}+[b_{ik}]_+b_{kj}+b_{ik}[-b_{kj}]_+ & \text{otherwise.} \end{cases} \]
  One easily observes that $\mu_k B$ is again skew-symmetrizable using the same matrix $D$.
  \begin{remark}
    Note that $\mu_k B=E_{k,\varepsilon} B F_{k,\varepsilon}$ for $\varepsilon=\pm 1$, the case $\varepsilon=1$ being obvious from the definitions and the case $\varepsilon=-1$ following from the identity $b_{ij}=[b_{ij}]_+-[-b_{ij}]_+$.
  \end{remark}

  To record sequences of these matrix mutations, we introduce the labeled $n$-regular rooted tree $\TT_n$ with root vertex $t_0$ and associate skew-symmetrizable matrices $B^t$ for $t\in\TT_n$ satisfying:
  \begin{itemize}
    \item $B^{t_0}=B$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $B^{t'}=\mu_k B^t$.
  \end{itemize}

  For $t\in\TT_n$, define $\phi^t_k:\RR^n\to\RR^n$ as the piecewise-linear map
  \[\phi^t_k(\lambda)=\begin{cases} E^t_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E^t_{k,-}\lambda & \text{if $\lambda_k<0$;} \end{cases}\]
  where the entries of $E^t_{k,\varepsilon}$ are given by \eqref{eq:left mutation matrix} with $b^t_{ij}$ in place of $b_{ij}$.
  We leave it as an exercise for the reader to check that $(\phi^t_k)^{-1}=\phi^{t'}_k$ whenever $t,t'\in\TT_n$ are joined by an edge labeled $k$.
  \begin{lemma}
    Suppose $t,t'\in\TT_n$ are joined by an edge labeled $k$.
    If $\mu\in\lambda+\Col(B^t)$, then $\phi^t_k(\mu)\in\phi^t_k(\lambda)+\Col(B^{t'})$.
  \end{lemma}
  \begin{proof}
    Suppose $\mu=\lambda+B^t\alpha$ for some $\alpha\in\RR^n$.
    Let $\varepsilon$ (resp. $\varepsilon'$) be such that $\phi^t_k(\mu)=E^t_{k,\varepsilon}\mu$ (resp. $\phi^t_k(\lambda)=E^t_{k,\varepsilon'}\lambda$).
    Then 
    \[\phi^t_k(\mu)=E^t_{k,\varepsilon}\mu=E^t_{k,\varepsilon}(\lambda+B^t\alpha)=E^t_{k,\varepsilon}\lambda + E^t_{k,\varepsilon} B^t F^t_{k,\varepsilon} F^t_{k,\varepsilon}\alpha=E^t_{k,\varepsilon}\lambda + B^{t'} F^t_{k,\varepsilon}\alpha.\]
    If $\varepsilon'=\varepsilon$, this is equal to $\phi^t_k(\lambda) + B^{t'} F^t_{k,\varepsilon}\alpha$ and so $\phi^t_k(\mu)\in\phi^t_k(\lambda)+\Col(B^{t'})$.
    Otherwise, 
    \[E^t_{k,\varepsilon}\lambda=E^t_{k,\varepsilon'}\lambda+\varepsilon(B^t)^{\bullet k}\lambda=\phi^t_k(\lambda)-\varepsilon(B^{t'})^{\bullet k}\lambda=\phi^t_k(\lambda)-\varepsilon B^{t'} \lambda_k \mathbf{e}_k\]
    and again we see that $\phi^t_k(\mu)\in\phi^t_k(\lambda)+\Col(B^{t'})$.
  \end{proof}

  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}

  Suppose $B$ is acyclic.
  Given a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, define a quiver $Q_{B,\varepsilon}$ from $Q_B$ by removing all arrows starting at vertex $i$ whenever $\varepsilon_i=-1$.
  For $1\le j\le n$, define $\cP_j(Q_{B,\varepsilon})$ as the set of paths in $Q_{B,\varepsilon}$ starting at vertex $j$.
  Elements of $\cP_j(Q_{B,\varepsilon})$ are sequences of vertices $(p_1,\ldots,p_r)$ with $p_1=j$ and $b_{p_{t+1}p_t}>0$ for $1\le t\le r-1$.
  Given a path $p\in\cP_j(Q_{B,\varepsilon})$, a vertex $i\in Q_0$, and a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, set
  \[\wt_{i,\varepsilon}(p):=-b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}\prod_{t=1}^{r-1} b_{p_{t+1}p_t},\]
  where the prefactor $b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}$ is omitted if $i=p_r$ or $\varepsilon_{p_r}=-1$.
  \begin{lemma}
    Suppose $B$ is acyclic with source-adapted sequence $k_1,\ldots,k_n$.
    Consider $t_1,\ldots,t_n\in\TT_n$ so that $t_{\ell-1}$ is joined to $t_\ell$ by an edge in $\TT_n$ labeled by $k_\ell$ for $1\le\ell\le n$.
    For a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, the product $E^{t_{n-1}}_{k_n,\varepsilon_{k_n}}\cdots E^{t_0}_{k_1,\varepsilon_{k_1}}$ is given by $\tau=(\tau_{ij})$ with
    \begin{equation}
      \label{eq:tropical twist}
      \tau_{ij}=\sum_{p\in\cP_j(Q_{B,\varepsilon})}^* \wt_{i,\varepsilon}(p),
    \end{equation}
    where the $*$ indicates that the summation includes the trivial path at $j$ only if $i\le j$ when $\varepsilon_j=1$ and only if $i=j$ when $\varepsilon_j=-1$.

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $n\times n$ matrix.
    Under the left action of $A$, say $A\bfx=\bfy$ for $\bfx,\bfy\in\RR^n$, the region inside $\RR^n$ defined by the inequality $\bfr\bullet(\bfx-\lambda)\le t$ with $\lambda,\bfr\in\RR^n$ and $t\in\RR$ is transformed into the region defined by the inequality \[(A^{-T}\bfr)\bullet(\bfy-A\lambda)\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\bfr\bullet(\bfx-\lambda)=\bfr^T(\bfx-\lambda)=\bfr^TA^{-1}A(\bfx-\lambda)=(A^{-T}\bfr)^T(A\bfx-A\lambda)=(A^{-T}\bfr)\bullet(\bfy-A\lambda).\]
  \end{proof}

>>>>

  The matrix $B$ is \emph{acyclic} if there is no sequence $i_1,\ldots,i_m,i_{m+1}=i_1$ so that $b_{i_\ell i_{\ell+1}}>0$ for $1\le\ell\le m$.
  This condition is naturally interpreted in terms of an associated quiver $Q_B$ with $n$ vertices and an arrow $i\to j$ whenever $b_{ji}>0$ (if we need to use representation theory at any point we'll have to change this to be $\gcd(b_{ij},b_{ji})$ arrows $i\to j$).
  Indeed, this is precisely the condition that $Q_B$ has no oriented cycles.
  In the case when $B$ is acyclic, there exists a labeling $\{k_1,\ldots,k_n\}=\{1,\ldots,n\}$ so that $\ell<\ell'$ implies $b_{k_{\ell'} k_\ell}\ge 0$.
  The sequence $(k_1,\ldots,k_n)$ (resp. $(k_n,\ldots,k_1)$) is then said to be \emph{source-adapted} (resp. \emph{sink-adapted}) since the condition implies $k_1$ is a source and $k_n$ is a sink in $Q_B$.

  We work with principal coefficients.
  Write $\tilde{B}=(b_{ij})$ for the $2n\times n$ matrix obtained from $B$ by adjoining an $n\times n$ identity matrix as the last $n$ rows.
  Call the matrix $B$ the \emph{principal part} of $\tilde{B}$.

  For $b\in\ZZ$, write $[b]_+=max(b,0)$.
  Given a sign $\varepsilon\in\{\pm1\}$ and $k\in\{1,\ldots,n\}$, define a $2n\times 2n$ matrix $E_{k,\varepsilon}=(e_{ij})$ with
  \begin{equation}
    \label{eq:left mutation matrix}
    e_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ [\varepsilon b_{ik}]_+ & \text{if $i\ne j=k$;}\\ 0 & \text{otherwise;} \end{cases}
  \end{equation}
  and an $n\times n$ matrix $F_{k,\varepsilon}=(f_{ij})$ with
  \begin{equation}
    \label{eq:right mutation matrix}
    f_{ij}=\begin{cases} 1 & \text{if $k\ne i=j$;}\\ -1 & \text{if $k=i=j$;}\\ [-\varepsilon b_{kj}]_+ & \text{if $k=i\ne j$;}\\ 0 & \text{otherwise.} \end{cases}
  \end{equation}
  Observe that $E^2_{k,\varepsilon}=\Id$ and $F^2_{k,\varepsilon}=\Id$ for any $k$ and any choice of $\varepsilon$.
  \begin{lemma}
    For $\varepsilon\in\{\pm1\}$ and $k\in\{1,\ldots,n\}$, we have $E_{k,-\varepsilon}E_{k,\varepsilon}=I_n+\varepsilon B^{\bullet k}$ and $E_{k,\varepsilon}-E_{k,-\varepsilon}=\varepsilon B^{\bullet k}$.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $\varepsilon b_{ij}=[\varepsilon b_{ij}]_+-[-\varepsilon b_{ij}]_+$.
  \end{proof}

  The index $k\in\{1,\ldots,n\}$ also determines a new matrix $\mu_k \tilde{B}=(b'_{ij})$ given by
  \[ b'_{ij}=\begin{cases} -b_{ij} & \text{if $i=k$ or $j=k$;}\\ b_{ij}+[b_{ik}]_+b_{kj}+b_{ik}[-b_{kj}]_+ & \text{otherwise.} \end{cases} \]
    One easily observes that the principal part of $\mu_k \tilde{B}$ is again skew-symmetrizable using the same matrix $D$.
  \begin{remark}
    Note that $\mu_k \tilde{B}=E_{k,\varepsilon} \tilde{B} F_{k,\varepsilon}$ for $\varepsilon=\pm 1$, the case $\varepsilon=1$ being obvious from the definitions and the case $\varepsilon=-1$ following from the identity $b_{ij}=[b_{ij}]_+-[-b_{ij}]_+$.
  \end{remark}

  To record sequences of these matrix mutations, we introduce the labeled $n$-regular rooted tree $\TT_n$ with root vertex $t_0$ and associate $2n\times n$ matrices $\tilde{B}^t$ for $t\in\TT_n$ satisfying:
  \begin{itemize}
    \item $\tilde{B}^{t_0}=\tilde{B}$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\tilde{B}^{t'}=\mu_k \tilde{B}^t$.
  \end{itemize}

  For $t\in\TT_n$ and $k\in\{1,\ldots,n\}$, define $\phi^t_k:\RR^{2n}\to\RR^{2n}$ as the piecewise-linear map
  \[\phi^t_k(\lambda)=\begin{cases} E^t_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E^t_{k,-}\lambda & \text{if $\lambda_k<0$;} \end{cases}\]
  where the entries of $E^t_{k,\varepsilon}$ are given by \eqref{eq:left mutation matrix} with $b^t_{ij}$ in place of $b_{ij}$.
  We leave it as an exercise for the reader to check that $(\phi^t_k)^{-1}=\phi^{t'}_k$ whenever $t,t'\in\TT_n$ are joined by an edge labeled $k$.
  \begin{lemma}
    Suppose $t,t'\in\TT_n$ are joined by an edge labeled $k$.
    If $\mu\in\lambda+\Col(\tilde{B}^t)$, then $\phi^t_k(\mu)\in\phi^t_k(\lambda)+\Col(\tilde{B}^{t'})$.
  \end{lemma}
  \begin{proof}
    Suppose $\mu=\lambda+\tilde{B}^t\alpha$ for some $\alpha\in\RR^n$.
    Let $\varepsilon$ (resp. $\varepsilon'$) be such that $\phi^t_k(\mu)=E^t_{k,\varepsilon}\mu$ (resp. $\phi^t_k(\lambda)=E^t_{k,\varepsilon'}\lambda$).
    Then 
    \[\phi^t_k(\mu)=E^t_{k,\varepsilon}\mu=E^t_{k,\varepsilon}(\lambda+\tilde{B}^t\alpha)=E^t_{k,\varepsilon}\lambda + E^t_{k,\varepsilon} \tilde{B}^t F^t_{k,\varepsilon} F^t_{k,\varepsilon}\alpha=E^t_{k,\varepsilon}\lambda + \tilde{B}^{t'} F^t_{k,\varepsilon}\alpha.\]
    If $\varepsilon'=\varepsilon$, this is equal to $\phi^t_k(\lambda) + \tilde{B}^{t'} F^t_{k,\varepsilon}\alpha$ and so $\phi^t_k(\mu)\in\phi^t_k(\lambda)+\Col(\tilde{B}^{t'})$.
    Otherwise, 
    \[E^t_{k,\varepsilon}\lambda=E^t_{k,\varepsilon'}\lambda+\varepsilon(\tilde{B}^t)^{\bullet k}\lambda=\phi^t_k(\lambda)-\varepsilon(\tilde{B}^{t'})^{\bullet k}\lambda=\phi^t_k(\lambda)-\varepsilon \tilde{B}^{t'} \lambda_k \mathbf{e}_k\]
    and again we see that $\phi^t_k(\mu)\in\phi^t_k(\lambda)+\Col(\tilde{B}^{t'})$.
  \end{proof}

  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^{2n}\to\RR^{2n}$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}

  Suppose the principal part $B$ of $\tilde{B}$ is acyclic.
  Given a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, define a quiver $Q_{B,\varepsilon}$ from $Q_B$ by removing all arrows starting at vertex $i$ whenever $\varepsilon_i=-1$.
  For $1\le j\le n$, define $\cP_j(Q_{B,\varepsilon})$ as the set of paths in $Q_{B,\varepsilon}$ starting at vertex $j$.
  Elements of $\cP_j(Q_{B,\varepsilon})$ are sequences of vertices $(p_1,\ldots,p_r)$ with $p_1=j$ and $b_{p_{t+1}p_t}>0$ for $1\le t\le r-1$.
  Given a path $p\in\cP_j(Q_{B,\varepsilon})$, a vertex $i\in Q_0$, and a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, set
  \[\wt_{i,\varepsilon}(p):=-b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}\prod_{t=1}^{r-1} b_{p_{t+1}p_t},\]
  where the prefactor $b_{ip_r}^{(1-\delta_{ip_r})(1+\varepsilon_{p_r})/2}$ is omitted if $i=p_r$ or $\varepsilon_{p_r}=-1$.
  \begin{lemma}
    Suppose $B$ is acyclic with source-adapted sequence $k_1,\ldots,k_n$.
    Consider $t_1,\ldots,t_n\in\TT_n$ so that $t_{\ell-1}$ is joined to $t_\ell$ by an edge in $\TT_n$ labeled by $k_\ell$ for $1\le\ell\le n$.
    For a sequence of signs $\varepsilon=(\varepsilon_1,\ldots,\varepsilon_n)$, the product $E^{t_{n-1}}_{k_n,\varepsilon_{k_n}}\cdots E^{t_0}_{k_1,\varepsilon_{k_1}}$ is given by $\tau=(\tau_{ij})$ with
    \begin{equation}
      \label{eq:tropical twist}
      \tau_{ij}=\sum_{p\in\cP_j(Q_{B,\varepsilon})}^* \wt_{i,\varepsilon}(p),
    \end{equation}
    where the $*$ indicates that the summation includes the trivial path at $j$ only if $i\le j$ when $\varepsilon_j=1$ and only if $i=j$ when $\varepsilon_j=-1$.

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $n\times n$ matrix.
    Under the left action of $A$, say $A\bfx=\bfy$ for $\bfx,\bfy\in\RR^n$, the region inside $\RR^n$ defined by the inequality $\langle\bfr,\bfx-\lambda\rangle\le t$ with $\lambda,\bfr\in\RR^n$ and $t\in\RR$ is transformed into the region defined by the inequality \[\langle A^{-T}\bfr,\bfy-A\lambda\rangle\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\langle\bfr,\bfx-\lambda\rangle=\bfr^T(\bfx-\lambda)=\bfr^TA^{-1}A(\bfx-\lambda)=(A^{-T}\bfr)^T(A\bfx-A\lambda)=\langle A^{-T}\bfr,\bfy-A\lambda\rangle.\]
  \end{proof}
